Predicting Remaining Useful Life of NASA Aircraft Engines 

### 2025-02-15 index.qmd
Updated Intro with html
Added link to article summaries
Updated Intro Take 2

---
title: "Cox Proportional Hazards Model"
subtitle: "Temp"
author: "Jayme Reed & Brad Paton (Advisor: Dr. Cohen)"
date: '`r Sys.Date()`'
format:
  html:
    code-fold: true
course: Capstone Projects in Data Science
bibliography: references.bib # file contains bibtex for references
#always_allow_html: true # this allows to get PDF with HTML features
self-contained: true
execute: 
  warning: false
  message: false
editor: 
  markdown: 
    wrap: 72
---

@@

## Introduction

The introduction should:

@@

Analyzing survival data is challenging for traditional machine learning models because they struggle to capture the time-to-event nature of the data.Survival data typically contains challenges difficult to account for because the datasets may be “censored” meaning that the event hasn’t had the opportunity to occur yet by the end of the study period. Also, traditional methods aren’t able to effectively model time-to-event data because they treat time as fixed values opposed to treating the values as time leading to less accurate results.

>>>>

---
title: "Predicting Remaining Useful Life of NASA Aircraft Engines "
subtitle: "Using Cox Proportional Hazards Model"
author: "Jayme Reed & Brad Paton (Advisor: Dr. Cohen)"
date: '`r Sys.Date()`'
format:
  html:
    code-fold: true
course: Capstone Projects in Data Science
bibliography: references.bib # file contains bibtex for references
#always_allow_html: true # this allows to get PDF with HTML features
self-contained: true
execute: 
  warning: false
  message: false
editor: 
  markdown: 
    wrap: 72
---

@@

[Literature](articleSummaries.html)

## Introduction (take 2)

As the COVID-19 pandemic demonstrated, the ability to predict the
survival of a patient based on their symptoms and medical history is a
major benefit as society deals with health crisis like pandemics. Those
predictions would assist medical practitioners in triage and allow
researchers the ability to generate predictions on the individuals who
are at the greatest risk of losing their lives due to an illness. While
the medical world is the most obvious use of survival predictions, the
same methods can be applied to the predicting bank failure, the
remaining useful life of machines, machinery maintenance timelines, and
insurance likelihood of payouts. There are various survival methods that
can be examined, but this paper will focus on Cox proportional hazards
model. (*don't like)*

Cox proportional hazards (CPH) model is a statistical regression model
for survival modeling that specializes in time-to-event predictions
using using single or multiple covariates and the effects they have on
survival. Traditional regression models struggle to capture the
time-to-event nature of data and treat the time variable as a fixed
value instead of continuous variable. CPH models are able to overcome
those obstacles and are capable of dealing with survival data that
contains censored (events that have not occurred yet in the study
period) data as well. The models also have the advantage of being
semi-parametric which allows more flexibility as compared to other
models.

Analyzing survival data is challenging for traditional machine learning
models because they struggle to capture the time-to-event nature of the
data. Survival data typically contains challenges difficult to account
for because the datasets may be “censored” meaning that the event hasn’t
had the opportunity to occur yet by the end of the study period. Also,
traditional methods aren’t able to effectively model time-to-event data
because they treat time as fixed values opposed to treating the values
as time leading to less accurate results.

*Intro guidelines:* The introduction should:

@@
